---
title: "The Normal Distribution"
output:
  ioslides_presentation:
    incremental: true
---

```{r setup, include=FALSE}
library(knitr)
options(digits=2)
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```


## Small n {.build}

Let $X$ be the total number of people in a sample of size 3 that favor same-sex
marriage if the true population proportion that favor is 0.5.

$$ X \sim Bin(n = 3, p = .5) \, \textrm{vs} \, X \sim N(\mu = np, \sigma = \sqrt{np(1 - p))}$$

```{r, echo=FALSE, eval=TRUE, fig.height=3.7}
x1  <- 0:3
df <- data.frame(x = x1, y = dbinom(x1, 3, 0.5))
qplot(x, y, data = df, ylim = c(0, .6), xlim = c(-1, 4)) + 
  geom_bar(stat = "identity", col = "darkgreen", fill = "white") +
  stat_function(fun = dnorm, args = list(mean = 1.5, sd = sqrt(3/4)), 
                col = "goldenrod", lwd = 1.2)
```


## Large n {.build}

Let $X$ be the total number of people in a sample of size 75 that favor same-sex
marriage if the true population proportion that favor is 0.5.

$$ X \sim Binom(n = 75, p = .5) \, \textrm{vs} \, X \sim N(\mu = np, \sigma = \sqrt{np(1 - p)})$$

```{r, echo=FALSE, eval=TRUE, fig.height=3.7}
x1  <- 0:75
df <- data.frame(x = x1, y = dbinom(x1, 75, 0.5))
qplot(x, y, data = df) + 
  geom_bar(stat = "identity", col = "darkgreen", fill = "white") +
  stat_function(fun = dnorm, args = list(mean = 75/2, sd = sqrt(74/4)), 
                col = "goldenrod", lwd = 1.2)
#(1 - pbinom(50, 75, .5)) * 2
```


## How large is large enough? {.build}

Rule of thumb:

$$ np ≥ 10 \\
\textrm{and} \\
n(1 − p) ≥ 10 $$

The number of expected successes and failures are each greater than 10.


## Why use the Normal at all? {.build}

- Simulation and exact Binomial can be computationally expensive.
- Normal is sometimes more generally applicable.
- Normal is fast.




